### 2024-06-24 benchmark.tex

\section{CNRS Benchmark}

As identified in Section \ref{sec:msfr-critique},
Moltres will benefit from further \gls{VV} of its
\gls{MSR} modeling capabilities. This chapter presents verification
results from Moltres for problems within the CNRS Benchmark
\cite{tiberga_results_2020}---a numerical benchmark specifically designed to
assess \gls{MSR} simulation tools on coupled multiphysics simulations of
fast-spectrum \glspl{MSR}. The benchmark consists of several steps in which
steady-state and transient simulations are prescribed. The benchmark starts with simple
single-physics cases and incrementally introduces various types of
multiphysics coupling until full coupling is simulated in the final steps. This
gradual approach helps participants identify sources of discrepancies
that can arise from differences in modeling assumptions or
cross-section libraries, or actual errors in the software. I adapted the
contents of this chapter from my journal article titled
``\textit{Verification of Moltres for Multiphysics Simulations of Fast-Spectrum
Molten Salt Reactors}'' and published in the Annals of Nuclear Energy journal
\cite{park_verification_2022}.

\input{cnrs-benchmark/benchmark}
\input{cnrs-benchmark/moltres}
\input{cnrs-benchmark/results}
\input{cnrs-benchmark/conclusion}