### 2024-08-11 msre.tex

\section{Group Constant Data Generation on OpenMC}

The group constants required by the neutron diffusion or $S_N$ neutron transport
methods are:
%
\begin{itemize}
  \item $\Sigma_{t,g}$: Macroscopic total cross section in group $g$,
  \item $\Sigma_{r,g}$: Macroscopic removal cross section in group $g$,
  \item $\Sigma_s^{g'\rightarrow g}$: Macroscopic group-to-group scattering cross section matrix,
  \item $\Sigma_{s,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering cross section matrix,
  \item $\Sigma_{sp,l}^{g'\rightarrow g}$: $l$-th Legendre moment of the macroscopic
    group-to-group scattering production cross section matrix,
  \item $D_g$: $P_1$-based diffusion coefficient in group $g$,
  \item $\nu\Sigma_{f,g}$: Product of the average number of neutrons produced per fission and the
    macroscopic fission cross section in group $g$,
  \item $\chi_g$: Neutron fission spectrum in group $g$.
\end{itemize}
%
I generated the group constants using OpenMC's multigroup cross section generation capability
\cite{boyd_multigroup_2019}. The group constant postprocessing script for Moltres
(\texttt{moltres\_xs.py}) calculates $\Sigma_{r,g}$ as
%
\begin{align}
  \Sigma_{r,g} =& \sum^G_{g'\neq g}\Sigma_s^{g\rightarrow g'}+\Sigma_{a,g}-\left(\Sigma_{sp}^{g
    \rightarrow g} - \Sigma_s^{g\rightarrow g}\right)
  \shortintertext{where}
      \Sigma_{a,g} =& \mbox{ macroscopic absorption cross section in group $g$,} \nonumber \\
      \Sigma_{sp}^{g\rightarrow g} =& \mbox{ macroscopic scattering production cross section from
      group $g$ to $g$.} \nonumber
\end{align}
%
$\Sigma_{r,g}$ primarily represents the loss of neutrons from group $g$ through outscattering and
absorption. $\Sigma_{sp}^{g\rightarrow g}$ incorporates neutron multiplication effects from neutron
knockout reactions into the scattering cross section. Neutron knockout reactions are commonly
tallied as scattering reactions, but $\Sigma_{r,g}$ is a convenient parameter in which to
incorporate neutron knockout effects in the neutron diffusion equations. I ran all test cases for
$S_N$ and hybrid method calculations with up to the 2nd Legendre moments of the scattering
cross sections ($L=2$).

OpenMC uses the $P_1$ flux-limited formulation \cite{pomraning_flux-limited_1984} for calculating
$D_g$ as follows
%
\begin{align}
  D_g =& \frac{1}{3\Sigma_{tr,g}},
  \shortintertext{where}
  \Sigma_{tr,g} =& \frac{\langle\Sigma_{t,g}\phi_g\rangle-\langle\Sigma_{s1,g}\phi_g\rangle}
  {\langle\phi_g\rangle}, \nonumber \\
  \langle\Sigma_{t,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\
  \Sigma_{t,g}(r,E)\Psi(r,E,\Omega), \nonumber \\
  \langle\Sigma_{s1,g}\phi_g\rangle =& \int_{r\in V}dr \int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE
  \int_{4\pi}d\Omega'\int^{\infty}_0dE'\int^1_{-1}d\mu\ \mu\Sigma_s(r,E'\rightarrow E,\Omega'\cdot
  \Omega)\phi(r,E',\Omega'), \nonumber \\
  \langle \phi \rangle =& \int_{r\in V}dr\int_{4\pi}d\Omega\int^{E_{g-1}}_{E_g}dE\ \Psi(r,E,\Omega)
  .\nonumber
\end{align}

I ran all \gls{MSRE} neutronics simulations with the eight neutron energy group structure proposed
by Jaradat for \gls{MSRE} analysis \cite{jaradat_development_2021-1}.
Table \ref{table:energy-group} shows the upper neutron energy bounds of the eight-group structure.
I selected this energy group structure as a compromise between accuracy and computational expense
among the various group structures investigated by Jaradat ranging from four to sixteen groups.

\begin{table}[htb]
  \centering
  \caption{Neutron energy group structure in this work. Originally devised by Jaradat
  \cite{jaradat_development_2021-1}.}
  \begin{tabular}{r S}
    \toprule
    Group & {Upper energy bound [eV]} \\
    \midrule
    1 & 2.000$\times 10^7$ \\
    2 & 1.353$\times 10^6$ \\
    3 & 6.734$\times 10^4$ \\
    4 & 9.118$\times 10^3$ \\
    5 & 1.487$\times 10^2$ \\
    6 & 4.000$\times 10^0$ \\
    7 & 6.250$\times 10^{-1}$ \\
    8 & 8.000$\times 10^{-2}$ \\
    \bottomrule
  \end{tabular}
  \label{table:energy-group}
\end{table}

I set the material compositions and densities of every component in the \gls{MSRE} models using
reference data from \gls{MSRE} benchmark specifications collated by Fratoni et al.
\cite{fratoni_molten_2020}. All OpenMC models utilize the ENDF/B-VII.1 nuclear data library
\cite{chadwick_endf/b-vii.1_2011} and include thermal scattering cross section data for beryllium
and carbon target nuclei in the fuel salt and graphite moderator. All models have a uniform
temperature of 900 K unless otherwise stated. Detailed information pertaining specifically to the
1-D, 2-D, and 3-D OpenMC models may be found in the respective subsections below.

\section{Neutronics Methods}

I ran all cases on OpenMC to generate the required group constants and reference $k_\text{eff}$
values to assess the accuracy of the deterministic methods against OpenMC in continuous-energy
(OpenMC-CE) and multigroup (OpenMC-MG) modes. Table \ref{table:var} shows how the
position, direction of travel, neutron energy, and angle-dependence in $\Sigma_s$ are represented
by OpenMC and the deterministic methods. The hybrid $S_N$-diffusion method features representations
from either $S_N$ or neutron diffusion depending on the subregion under consideration.
Comparing OpenMC-MG results with OpenMC-CE results allows us
to quantify errors arising from neutron energy group discretization and the scattering cross
section simplifications.

\begin{table}[htb!]
  \centering
  \footnotesize
  \caption{Variable representations in OpenMC under continuous-energy (OpenMC-CE) and multigroup
  (OpenMC-MG) modes, and in the $S_N$ neutron transport and neutron diffusion.}
  \begin{tabular}{l l l l l}
    \toprule
    Variable & OpenMC-CE & OpenMC-MG & $S_N$ Transport & Diffusion \\
    \midrule
    Position, $\bm{r}$ & Continuous & Continuous & Discrete & Discrete \\
    Direction of travel, $\bm{\hat{\Omega}}$ & Continuous & Continuous & Discrete & N/A \\
    Energy, $E$ & Continuous & Discrete & Discrete & Discrete \\
    Angle-dependence in $\Sigma_s$ & Continuous & $L=2$ & $L=2$ & N/A \\
    \bottomrule
  \end{tabular}
  \label{table:var}
\end{table}

The rest of this section pertains to the deterministic neutronics methods used in this work for
modeling the \gls{MSRE}.
For the 1-D models, I performed neutronics calculations with the $S_N$, neutron diffusion, and
hybrid method implementations on Moltres and Python. Table \ref{table:nts-methods} highlights the
differences between the two implementations. The $S_N$ neutron transport method implementation on
Moltres follows the methodology for solving the \gls{SAAF} formulation of the $S_N$ equations as
described in Section \label{sec:saaf} while the implementation on Python follows the methodology
for solving the original first-order formulation as described in Section \label{sec:python-sn}.

The implementations on Moltres are discretized in space using the finite element method
while the implementations on Python use finite difference or diamond difference schemes. The most
crucial difference lies in the transport correction formulations for the hybrid method. The Moltres
implementation is based on the drift correction formulation described in Section
\label{sec:drift-correction} while the Python implementation is based on the diffusion correction
formulation described in Section \label{sec:diffusion-correction}.

I will refer to the methods and results from either Molres or Python implementations with the
Moltres- or Python- prefix to clearly set them apart, i.e. Moltres-diffusion, Moltres-$S_N$,
Moltres-hybrid, Python-diffusion, Python-$S_N$, and Python-hybrid.

For 2- and 3-D models, I abandoned the diffusion correction formulation in favor of the drift
correction formulation due to superior numerical performance in the latter. The analysis
supporting this decision may be found in the analysis of 1-D results in Section
\ref{sec:1d-results}.

\begin{table}[htb!]
  \centering
  \footnotesize
  \caption{Implementation differences of neutronics methods on Moltres and Python.}
  \begin{tabular}{p{0.14\linewidth} p{0.25\linewidth} p{0.15\linewidth} p{0.15\linewidth} p{0.17\linewidth}}
    \toprule
    Implementations & Spatial discretization & $S_N$ formulation & $S_N$ acceleration &
    Hybrid transport\newline correction \\
    \midrule
    Moltres & Finite element method & SAAF & Nonlinear diffusion acceleration & Drift correction \\
    Python & Finite difference (Diffusion)\newline Diamond difference ($S_N$) & First-order & None & Diffusion correction \\
    \bottomrule
  \end{tabular}
  \label{table:nts-methods}
\end{table}

\section{Test Metrics} \label{sec:test-metrics}

In this section, I define several metrics for comparing simulation results among the various
neutronics methods.

For steady-state eigenvalue calculations, all solvers directly provide the effective neutron
multiplication factor ($k_\text{eff}$), the rate of neutron population increase between successive
neutron generations. Reactivity is defined as
%
\begin{gather}
  \rho \equiv \frac{k_\text{eff}-1}{k_\text{eff}}.
\end{gather}
%
Differences in reactivity, for either comparisons between neutronics methods or control rod worth
calculations, can be calculated as
%
\begin{gather}
  \Delta\rho_{1,2} = \rho_1 - \rho_2 =
  \frac{k_{\text{eff},1}-k_{\text{eff},2}}{k_{\text{eff},1}k_{\text{eff},2}}.
\end{gather}

Spatially dependent variables (e.g., neutron group flux, neutron drift) are sampled at fixed 0.1 cm
intervals along the specified lines (e.g., horizontal, diagonal, etc.) regardless of the actual
mesh resolution.

\section{1-D Models \& Results} \label{sec:1d-results}

\subsection{1-D Model Setup \& Geometries}

I designed six 1-D test cases with increasing complexity to verify the $S_N$ and hybrid
$S_N$-diffusion implementations and to test the performance of the hybrid
method in response to various geometrical features. The last two cases resemble the
reference \gls{MSRE} design \cite{robertson_msre_1965}, which has centrally located control rods
and air-filled rod guide tubes. Figure \ref{fig:case-geom} shows the geometries of Cases 1a to
3b. All geometries have reflective boundary conditions at $x=0$ cm to reduce computational costs by
creating half-core or repeating unit cell models. Cases 1a and 1b are repeating unit
cell models with reflecting boundaries on the right-side boundaries. Cases 2a, 2b, 3a, and 3b
are 1-D half-core models with vacuum boundaries on the right-side boundaries.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.8\columnwidth]{case-geometry}
  \caption{Geometries of the 1-D test cases. The material labeled ``mixture'' represents a
    homogeneous mixture of fuel and graphite at a ratio of 22.5\%-77.5\% by volume. All geometries
    have reflective boundary conditions on the boundary at $x=0$ cm. The right-side boundaries are
    reflective for Cases 1a and 1b, and vacuum for Cases 2a, 2b, 3a, and 3b.}
  \label{fig:case-geom}
\end{figure}

Cases 1a and 1b are simple test cases containing neutron multiplying regions. Cases 1a and 1b serve
to verify basic neutronics phenomena modeling (e.g., fission, scattering, absorption) of the newly
implemented $S_N$ and neutron diffusion methods in Python and $S_N$ method in Moltres. I did not
apply the hybrid method on these cases. Cases 2a and 2b represent 1-D models of the \gls{MSRE} with
air-filled control rod guide tube regions, a homogenized fuel-graphite lattice regions, and outer
vessel regions. Case 2b additionally contains a 0.5cm-thick control rod region to contrast with
Case 2a for control rod worth calculations. Cases 3a and 3b retain the heterogeneous fuel-graphite
lattice geometry present in the \gls{MSRE}.

In my preliminary investigations, I discovered that using the original control rod material
composition (70 wt\% Gd$_2$O$_3$-30 wt\% Al$_2$O$_3$) in the 1-D models led to excessively large
control rod worth estimates of approximately 55000 pcm. The worth estimates are ten times larger
than the 5500 pcm expected from full rod insertion of all three rods in the 3-D models
\cite{fratoni_molten_2020}. Therefore, I greatly reduced the proportion of Gd$_2$O$_3$ in the
control rod to 0.35 wt\% for a fairer assessment of rod worth in the 1-D models. This change
brought the new rod worth estimates down to below 20000 pcm.

\subsection{1-D Model Mesh Convergence Tests}

I ran mesh convergence studies based on Case 3b for the neutron diffusion and $S_N$ neutron
transport methods. Given that the hybrid $S_N$-diffusion method comprises of a combination of those
methods, hybrid simulations are fully spatially resolved on meshes that are appropriately resolved
for both neutron diffusion and $S_N$ simulations.

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-k}
    \caption{Neutron diffusion method}
    \label{fig:diff-mesh-k}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{sn-mesh-convergence-k}
    \caption{$S_8$ neutron transport method}
    \label{fig:sn-mesh-k}
  \end{subfigure}
  \caption{Convergence of multiplication factor ($k_\text{eff}$) estimates for Case 3b across five
    levels of mesh refinement. Higher refinement levels correspond to smaller mesh element sizes.}
  \label{fig:mesh-k}
\end{figure}

Figures \ref{fig:diff-mesh-k} and \ref{fig:sn-mesh-k} show the convergence of $k_\text{eff}$
estimates observed for Case 3b simulations with neutron diffusion and $S_8$ neutron transport
methods. The $k_\text{eff}$ values plateau at 0.9713203 and 0.9813790 for the neutron diffusion and
$S_8$ methods, respectively. By mesh refinement level 2, both methods are within $5\times 10^{-7}$
range of the $k_\text{eff}$ value at level 4.

Figures \ref{fig:diff-mesh-flux} and \ref{fig:sn-mesh-flux} show the absolute differences in the
group flux distributions at mesh refinement levels 1 to 3 relative to the reference group flux
distribution at level 4 (maximum refinement). I omitted data from level 0 to improve the fidelity
of the plots as flux differences drop by one order of magnitude with each level of refinement.
With these results, I concluded that the mesh resolutaion at level 2 is a sufficient guideline for
the mesh resolutions used in the subsequent 1-D analyses. The absolute differences in both
$k_\text{eff}$ and group flux distributions at level 2 remain below $5\times10^{-7}$ relative to
level 4.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{diffusion-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the neutron diffusion method for
  mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
  refinement level 5.}
  \label{fig:diff-mesh-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{sn-mesh-convergence-flux}
  \caption{Absolute difference in group fluxes ($\phi_g$) from the $S_8$ neutron transport method
  for mesh refinement levels 1 to 4 relative to the reference group flux distributions at mesh
  refinement level 5.}
  \label{fig:sn-mesh-flux}
\end{figure}

\FloatBarrier

\subsection{1-D Numerical Results}

\subsubsection{Reactivity}

Figure \ref{fig:1d-rho} shows the difference in reactivities of OpenMC-MG, $S_8$,
neutron diffusion, and hybrid methods relative to OpenMC-CE. The standard deviations of
reactivity values from OpenMC-CE are approximately 40 pcm for Cases 1a and 1b and 60 pcm for the
rest as depicted by the blue highlighted area in Figure \ref{fig:1d-rho}. For Cases 1a and 1b, the
OpenMC-MG, $S_8$, and neutron diffusion methods show excellent agreement with OpenMC-CE as the
reactivity values fall within the one or two standard deviation range. The Case 1 results show that
the Moltres and Python implementations are consistent with each other given that they are spatially
well-resolved.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{rho}
  \caption{Difference in reactivity $\rho$ of all neutronics methods investigated relative
  to OpenMC-CE.}
  \label{fig:1d-rho}
\end{figure}

For all cases, the OpenMC-MG and the $S_8$ methods show consistent agreement with one another. 
While they deviate from OpenMC-CE by approximately 350 pcm for Cases 2a and 3a, they fall within
two standard deviations of OpenMC-CE for Cases 2b and 3b. When increasing the maximum Legendre
polynomial order to approximate the angular dependence in $\Sigma_s^{g'\rightarrow g}$ from
2nd-order to 3rd-order, the reactivity changed by only 1 pcm. Therefore, the reactivity
discrepancies are mainly attributed to the eight neutron energy group structure which remains the
only significant difference between OpenMC-CE and the multigroup methods.

The neutron diffusion and hybrid methods agree closely with one another for Cases 2a and 3a which
exclude the control rod region. This shows that the hybrid method provides similar $k_\text{eff}$
estimates to the neutron diffusion method in 1-D simulations without highly neutron-absorbing
regions. Compared to the OpenMC-MG and $S_8$ reactivity values, the neutron diffusion and hybrid
method reactivity values agree closer with the reference OpenMC-CE value. However, this is likely
due to favorable error cancellation; Figures \ref{fig:3a-flux} and \ref{fig:3a-flux-diff} show that
OpenMC-MG and the $S_8$ methods reproduce
the OpenMC-CE flux distribution more accurately than the neutron diffusion and hybrid methods.

In Cases 2b and 3b, the neutron diffusion method largely fails to accurately capture the
effect of introducing the control rod region as evidenced by the -1500 pcm and -1150 pcm
discrepancies relative to OpenMC-CE. The hybrid method fares better with -400 pcm and -300 pcm
discrepancies.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{worth}
  \caption{Percentage difference in rod worth for Cases 2 and 3 of all neutronics methods
  investigated relative to OpenMC-CE.}
  \label{fig:1d-worth}
\end{figure}

\subsubsection{Control rod worth}

Moving the discussion to control rod worths, figure \ref{fig:1d-worth} shows the percentage
difference in rod worths for Cases 2 and 3 of OpenMC-MG, $S_8$, neutron diffusion, and hybrid
methods relative to OpenMC-CE. I calculated the rod worth estimates from each method by taking the
difference in reactivity between the ``rod out'' (a) and ``rod in'' (b) configurations of Cases 2
and 3. All neutronics methods overestimate the rod worth relative to OpenMC-CE, resulting in
positive percentage difference values observed in the figure.
The neutron diffusion methods clearly show up as outliers with rod worth estimates in excess
of 8\%. Most of the remaining methods cluster around 2.5\% and 3\% percentage difference for Cases
2 and 3 with the exception of the Python-hybrid method for Case 3.

Given that the hybrid methods rely on transport corrections derived from the $S_N$ method, the
$S_8$ rod worth estimates serve as the
reference point for hybrid method verification. Additionally, all hybrid and $S_N$ simulations are
equally affected by errors arising from the multigroup approximation. While the Python-hybrid
method produces the lowest percentage difference relative to OpenMC-CE for Case 3, the
Python-hybrid method deviates from the $S_8$ ``multigroup'' reference rod worths more than the
Moltres-hybrid method. Nevertheless, both hybrid method implementations provide significant
improvements in rod worth estimates over the neutron diffusion method.

\subsubsection{Neutron flux distribution}

This discussion on the neutron flux distributions focuses on Cases 3a and 3b which most closely
resemble the actual \gls{MSRE} reactor geometry. Cases 3a and 3b resemble the \gls{MSRE} reactor
with the control rod withdrawn and inserted, respectively. Figures \ref{fig:3a-flux} and
\ref{fig:3b-flux} show the neutron group flux distributions for Cases 3a and 3b from OpenMC-CE and
OpenMC-MG. Differences between the two methods arise from the multigroup approximation, notably in
group 2 for Case 3a and groups 2, 6, 7, 8 for Case 3b. Future work aimed at finding a better
neutron energy group structure should at minimum focus on improving fidelity in the listed energy
groups.

For a fair evaluation of the deterministic multigroup methods without distortions from the
multigroup approximation, I chose to compare their flux distributions with the OpenMC-MG flux
distribution. Figures \ref{fig:3a-flux-diff} and \ref{fig:3b-flux-diff} show the absolute
difference in neutron group flux distributions of the Moltres-$S_8$, Moltres-diffusion,
Moltres-hybrid, and Python-hybrid methods relative to OpenMC-MG. The figures omit Python-$S_8$ and
Python-diffusion as their distributions are nearly identical to their Moltres-based counterparts.
The neutron diffusion and hybrid methods fare worse than the $S_8$ method at capturing the
oscillatory pattern in groups 1, 2, 5, 7, and 8 arising from the fuel-graphite lattice geometry
along $x=5$ cm to 10 cm. In both Cases 3a and 3b, the neutron diffusion method (yellow plot lines)
exhibits significant flux deviations in groups 1, 2, 7, and 8 near $x=0$ cm.

Both hybrid methods
provide significant improvements in neutron flux distributions compared to the neutron diffusion
method as shown by the generally smaller flux difference values particularly near $x=0$ cm. The
figures also show that Moltres-hybrid (green) provides more accurate flux estimates than
Python-hybrid (red) in all instances. The differences indicate deficiencies in the diffusion
correction scheme in Python-hybrid that lead to poorer performance compared to the drift
correction scheme in Moltres-hybrid.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux}
  \caption{Case 3a neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3a-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux}
  \caption{Case 3b neutron group flux distributions from OpenMC-CE and OpenMC-MG.}
  \label{fig:3b-flux}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3a-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3a from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3a-flux-diff}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-flux-diff}
  \caption{Absolute difference in neutron group flux distributions for Case 3b from Moltres-$S_8$,
  Moltres-diffusion, Moltres-hybrid, and Python-hybrid relative to OpenMC-MG.}
  \label{fig:3b-flux-diff}
\end{figure}

\FloatBarrier

\subsubsection{Drift \& diffusion correction distribution}

Figures \ref{fig:3b-drift} and \ref{fig:3b-svdc} show the distributions of the drift and diffusion
correction parameters, respectively, in each neutron energy group for Case 3b. As a reminder to the
reader, the hybrid $S_N$-diffusion methods generate these correction parameters in subregions
containing highly neutron absorbing materials. I implemented the drift correction scheme for
Moltres-hybrid and the diffusion correction scheme for Python-hybrid. The hybrid methods
selectively apply the $S_N$ method to improve neutronic modeling within the subregions. I computed
the reference distributions (yellow data points) in both plots from reference flux solutions of the
standard $S_8$ method.

As discussed in Section \ref{sec:buffer-region}, both sets of correction parameters generated by
the two hybrid methods match their respective reference values within the correction subregion from
$x=0$ cm to 10 cm except near the subregion boundary at $x=10$ cm. Therefore, both hybrid methods
provide improved flux estimates in the subregion compared to the neutron diffusion method.
The prior reactivity and flux distribution results indicate that Moltres-hybrid is more accurate
than Python-hybrid relative to OpenMC-MG as the reference.

In Figure \ref{fig:3b-drift}, the drift
correction distributions (Moltres-hybrid) for all eight energy groups are well-defined throughout
the subregion and are continuous except at material interfaces. Consequently, the drift correction
scheme provides accurate flux corrections for about 75\% of the subregion based on the buffer
region cutoff criteria (refer to Section \ref{sec:buffer-region}). By contrast, the diffusion
correction distributions in Figure \ref{fig:3b-svdc} contains numerous discontinuities tending towards
positive or negative infinity in all eight energy groups. The discontinuities arise when
computing corrected diffusion coefficients (Eq. \ref{eq:svdc}) near flux peaks and troughs
(stationary points) where the flux gradients approach zero. The location of stationary points from
the diffusion and $S_N$ calculations in the hybrid method do not necessarily coincide, resulting in
discontinuous flux gradients in the final flux solution. Discontinuities are generally unavoidable
in most problems because the fluxes typically reach local minima in control rod regions.
Furthermore, unlike negative drift coefficients, negative diffusion coefficients are unphysical.
Taking everything into consideration, the diffusion correction scheme is less accurate and more
numerically unstable than the drift correction scheme.

The subsequent discussions on the correction region size and the number of outer
iterations focus on the drift correction scheme of the hybrid method.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-drift}
  \caption{Multigroup drift correction ($\vec{D}_g$) $x$-component distributions from the
  Moltres-hybrid and Moltres-$S_8$ solvers.}
  \label{fig:3b-drift}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{case-3b-svdc}
  \caption{Multigroup diffusion correction ($D^s_g$) $x$-component distributions from the
    Python-hybrid and Python-$S_8$ solvers.}
  \label{fig:3b-svdc}
\end{figure}

\FloatBarrier

\subsubsection{Impact of correction subregion sizes}

The hybrid $S_N$-diffusion method relies on $S_N$ calculations in the correction subregion to
generate flux corrections. Minimizing the size of the correction subregion and the $S_N$ subproblem
is essential for the hybrid method to be computationally competitive for time-dependent full-core
simulations. I investigated the effect of the correction subregion size on the $k_\text{eff}$ and
control rod worth estimates with Cases 3a and 3b by varying the subregion sizes from 10 cm to 40 cm
at 5 cm-intervals.

Figures \ref{fig:v1-size-a-k} and \ref{fig:v1-size-b-k} show the $k_\text{eff}$ estimates from the
hybrid method for Cases 3a and 3b, respectively. In both cases, the $k_\text{eff}$ values initially
decrease as the correction subregion sizes increase before reversing in trend when the subregion
size reaches 35 cm and beyond. The $k_\text{eff}$ values vary by up to 164 pcm for Case 3a and 109
pcm for Case 3b. An important observation here is that the $k_\text{eff}$ values do not
monotonically converge towards the $k_\text{eff}$ estimate from the $S_8$ method. The data implies
that other significant sources of discrepancies exist beyond those being corrected in the
correction subregion by the hybrid method.

% In general,
% discrepancies arise between the neutron diffusion method and $S_8$ from the neutron diffusion 
% method's failure to accurately model the neutron flux in the fuel-graphite lattice region. The
% figures imply that those discrepancies result in a positive change in $k_\text{eff}$ since the
% $k_\text{eff}$ decreases as larger fractions of the problem domain benefit from flux corrections.
% The remaining discrepancy in $k_\text{eff}$ between the hybrid method with a correction subregion
% size of 40 cm and the $S_8$ method arises from

\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.49\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{correction-size-a-k}
    \caption{Case 3a}
    \label{fig:v1-size-a-k}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{correction-size-b-k}
    \caption{Case 3b}
    \label{fig:v1-size-b-k}
  \end{subfigure}
  \caption{$k_\text{eff}$ estimates from the hybrid method for Cases 3a and 3b with different
  correction subregion sizes. The horizontal lines indicate $k_\text{eff}$ estimates from the
  OpenMC-CE, OpenMC-MG, and $S_8$ methods.}
  \label{fig:v1-size-k}
\end{figure}

Figure \ref{fig:v1-size-rho} shows the percentage difference in rod worth relative to OpenMC-CE. 
Due to the identical trends observed in the $k_\text{eff}$ estimates of both cases, the control rod
worth estimates do not change significantly when the correction subregion
sizes change. This indicates that limiting the correction subregion size to save
computational cost on the expensive $S_N$ calculations has a negligible impact on rod worth
estimates. The rod worth estimates for all investigated correction subregion sizes remain within
0.2\% of the $S_8$ method.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.8\columnwidth]{correction-size-rho}
  \caption{Percentage difference in rod worth from the hybrid method relative to OpenMC-CE for
    Cases 3a and 3b with different correction subregion sizes. The horizontal lines indicate
    equivalent rod worth differences from the OpenMC-MG and $S_8$ methods.}
  \label{fig:v1-size-rho}
\end{figure}

The $k_\text{eff}$ and rod worth results show that the hybrid method is able to accurately
reproduce rod worth estimates as long as the correction subregion size remains consistent. The
correction subregion size must also be sufficiently large to capture the highly neutron-absorptive
rod's influence on the neutron flux around the rod at $x=0$ cm. The drift correction parameter
distributions in Figure \ref{fig:3b-drift} indicate that the influence of the rod on the drift, and
consequently the flux, extend to approximately $x=5$ cm before the drift distribution settles on a
regular repeating pattern influenced by the fuel-graphite lattice structure. I applied a consistent
correction subregion size of 10 cm for Cases 2 and 3 since it sufficiently meets the requirements
discussed while also minimizing the computational cost of the $S_8$ subsolver. Other reactor
designs consisting of different material components may require different minimum correction
subregion sizes.

\subsubsection{Relaxing the $S_N$ subsolver convergence threshold}

As described in Section \ref{sec:hybrid-algorithm}, the hybrid $S_N$-diffusion method works by
iteratively coupling a neutron diffusion solver and a $S_N$ subsolver with the latter generating
drift correction parameters for the former. Excluding the initial neutron diffusion calculation,
each outer (fixed point) iteration comprises of an inner $S_N$ calculation on the correction
subregion and an inner neutron diffusion calculation on the entire domain. The \gls{FEM} algorithm
in Moltres determines that an individual solver reaches convergence when the combined
residual contributions fall below a specified convergence tolerance threshold. The fixed point
iteration scheme under the \texttt{MultiApps} system in Moltres for coupling multiple individual
solvers (referred to as apps) requires users to specify a main app for the final convergence check
after each outer iteration to terminate the iterations. The scheme also requires users to set a
fixed point convergence tolerance value independent of the tolerance values of each individual
solver. Naturally, the fixed point tolerance value must be equal to or greater than the tolerance
value of the main individual solver for the possibility of overall convergence. The hybrid method
implementation in Moltres sets the neutron diffusion solver as the main app and the $S_N$ subsolver
as the sub app.

The hybrid method shares some similarities with diffusion-based \gls{HOLO} acceleration schemes
(refer to Section \ref{sec:drift-correction}) for the neutron transport methods which do not
require the neutron transport side of the calculations to fully converge
\cite{reynolds_analysis_2023, wang_diffusion_2014}.
The typical convergence threshold value required for converged neutron flux calculations in Moltres
is $10^{-8}$; I set the convergence tolerance value to $10^{-8}$ for the neutron diffusion solver
and the fixed point scheme. I varied the convergence tolerance value for the $S_N$ subsolver to 
find the optimal value that maintains sufficient solution accuracy and minimizes computational
costs; tightening the $S_N$ convergence tolerance value generally results in more outer iterations
performed before the solver meets the fixed point convergence tolerance.
Table \ref{table:sn-tol} lists the number of outer iterations required by the hybrid method for
a given set of $S_8$ subsolver convergence tolerance values in Case 3b.

\begin{table}[htb]
  \centering
  \caption{Number of outer iterations in hybrid method calculations of Case 3b for a given set of
  convergence tolerance values imposed on the $S_8$ subsolver.}
  \begin{tabular}{l S S S S S S}
    \toprule
    $S_8$ subsolver tolerance, $\epsilon_\text{tol}$ & {$10^{-8}$} & {$10^{-7}$} & {$10^{-6}$} & {$10^{-5}$} & {$10^{-4}$} & {$10^{-3}$} \\
    \midrule
    Number of outer iterations & 3 & 3 & 3 & 2 & 2 & 1 \\
    \bottomrule
  \end{tabular}
  \label{table:sn-tol}
\end{table}

Figure \ref{fig:sn-tol} shows the $k_\text{eff}$ error estimates relative to the reference value
when $\epsilon_\text{tol}=10^{-8}$. The hybrid method exhibits superlinear convergence ($q=1.333$)
with respect to the $S_N$ subsolver convergence tolerance value. Based on the results, setting
$\epsilon_\text{tol}$ to $10^{-5}$ is the best balance between accuracy and computational
cost because the $k_\text{eff}$ error is less than 0.01 pcm and the number of outer iterations
would increase from 2 to 3 if the tolerance value is further tightened.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.6\columnwidth]{sn-tol}
  \caption{$k_\text{eff}$ error estimates of Case 3b for a range of convergence tolerance values
  imposed on the $S_8$ subsolver relative to the reference $k_\text{eff}$ value when
  $\epsilon_\text{tol}=10^{-8}$. The $q=1.333$ line represents the approximate rate of
  convergence.}
  \label{fig:sn-tol}
\end{figure}

\section{2-D Models \& Results}

\subsection{2-D Model Setup}

Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
\gls{MSRE} models investigated. They are similar to the horizontal cross section of the base
\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
in the 2-D \gls{MSRE} model for this work:

\begin{itemize}
  \item Ignored thermal expansion of all reactor components.
  \item Modeled the control rods as perfectly annular cylinders instead of strings of beaded
    control elements. Ignored the inconel shells encasing the control elements and the steel
    helical cables linking the control elements together.
  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
  \item Homogenized the INOR, graphite, and molten salt regions in the sample basket.
  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
  \item Replaced partial channels with full-size channels near the edge of the core graphite.
\end{itemize}

The differences listed above, along with other changes in the 3-D model, result in negligible
changes to the $k\text{eff}$ (See Section 5.6) relative to the \gls{MSRE} benchmark model
\cite{fratoni_molten_2020}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{quarter-core-geom}
  \caption{2-D \gls{MSRE} quarter-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:1/4-geom}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{full-core-geom}
  \caption{2-D \gls{MSRE} full-core model based on the horizontal cross section of the actual
  \gls{MSRE} geometry.}
  \label{fig:full-geom}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.7\columnwidth]{full-core-closeup}
  \caption{Detailed view of the control rod thimbles and sample basket in 2-D \gls{MSRE} full-core
    model.}
  \label{fig:full-geom-closeup}
\end{figure}

\FloatBarrier

\subsection{2-D Numerical Results}

\subsubsection{2-D quarter-core results}

\begin{table}[htb]
  \small
  \centering
  \caption{$k_\text{eff}$ and control rod worth estimates for the 2-D quarter-core \gls{MSRE}
    model. Error estimates are relative to OpenMC-CE.}
  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=4(2)] S}
    \toprule
    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} & \multicolumn{2}{c}{Rod worth} \\
                            & {$k_\text{eff}$} & {Error [pcm]} & {$k_\text{eff}$} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
	  OpenMC-CE & 1.11209(43) & {-} & 1.01740(42) & {-} & 8370(53) & {-} \\
	  OpenMC-MG & 1.11979(42) & 618 & 1.02204(41) & 446 & 8541(51) & 172 \\
      Hybrid & 1.12174 & 773 & 1.02532 & 760 & 8383 & 13 \\
      Diffusion & 1.12059 & 682 & 1.00903 & -816 & 9867 & -1484 \\
    \bottomrule
  \end{tabular}
  \label{table:quarter-core}
\end{table}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-no-rod}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-rod}
\end{figure}

\begin{table}[htb]
  \small
  \centering
  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
  the 2-D \gls{MSRE} quarter-core models relative to OpenMC. The average standard deviation of
  OpenMC normalized channel fission rates is 0.2\%.}
  \begin{tabular}{l S S S S}
    \toprule
    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod} \\
                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-5}
    Hybrid & 0.40 & 1.32 & 0.43 & 3.08 \\
    Diffusion & 0.40 & 2.63 & 2.01 & 17.44 \\
    \bottomrule
  \end{tabular}
  \label{table:quarter-core-power}
\end{table}

\FloatBarrier

\subsubsection{2-D full-core results}

\begin{table}[htb]
  \small
  \centering
  \caption{$k_\text{eff}$ estimates for the 2-D full-core \gls{MSRE} model with the indicated rods
    inserted. Error estimates are relative to OpenMC-CE.}
  \setlength\tabcolsep{2pt}
  \begin{tabular}{l S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S S[table-format=1.5(2)] S}
    \toprule
    \multirow{3}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
                            & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} & {\multirow{2}{*}{$k_\text{eff}$}} & {Error} \\
                            & & {[pcm]} & & {[pcm]} & & {[pcm]} & & {[pcm]} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
    OpenMC-CE & 1.09859(20) & {-} & 1.06980(21) & {-} & 1.04690(18) & {-} & 1.02688(18) & {-} \\
    OpenMC-MG & 1.10637(19) & 640 & 1.07633(20) & 567 & 1.05235(20) & 495 & 1.03262(17) & 541 \\
    Hybrid & 1.10843 & 808 & 1.07906 & 802 & 1.05553 & 781 & 1.03582 & 841 \\
    Diffusion & 1.10701 & 692 & 1.07120 & 122 & 1.04415 & -252 & 1.02195 & -470 \\
    \bottomrule
  \end{tabular}
  \label{table:full-core-k}
\end{table}

\begin{table}[htb]
  \small
  \centering
  \caption{Control rod worth estimates for the 2-D full-core \gls{MSRE} with the
  indicated rods inserted. Error estimates are relative to OpenMC-CE.}
  \setlength\tabcolsep{5pt}
  \begin{tabular}{l S[table-format=4(2)] S S[table-format=4(2)] S S[table-format=4(2)] S}
    \toprule
    \multirow{2}{*}{Method} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
                            & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} & {$\Delta\rho_\text{worth}$ [pcm]} & {Error [pcm]} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(l){6-7}
    OpenMC-CE & 2450(25) & {-} & 4494(23) & {-} & 6357(24) & {-} \\
    OpenMC-MG & 2523(23) & 73 & 4640(24) & 146 & 6455(22) & 98 \\
    Hybrid & 2455 & 5 & 4521 & 27 & 6323 & -34 \\
    Diffusion & 3019 & 569 & 5439 & 945 & 7519 & 1162 \\
    \bottomrule
  \end{tabular}
  \label{table:full-core-worth}
\end{table}

\begin{table}[htb]
  \footnotesize
  \centering
  \caption{Absolute mean and maximum percentage errors in the normalized channel fission rates of
  the 2-D \gls{MSRE} full-core models relative to OpenMC. The average standard deviation of
  OpenMC normalized channel fission rates is 0.27\%.}
  \setlength\tabcolsep{2.5pt}
  \begin{tabular}{l S S S S S S S S}
    \toprule
    \multirow{2}{*}{Method} & \multicolumn{2}{c}{No Rod} & \multicolumn{2}{c}{Rod 1} & \multicolumn{2}{c}{Rod 1 \& 2} & \multicolumn{2}{c}{Rod 1, 2 \& 3} \\
                            & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} & {Mean [\%]} & {Maximum [\%]} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(rl){4-5} \cmidrule(rl){6-7} \cmidrule(l){8-9}
    Hybrid & 0.43 & 1.45 & 0.43 & 1.82 & 0.43 & 2.26 & 0.43 & 2.52 \\
    Diffusion & 0.45 & 2.95 & 0.94 & 12.61 & 1.35 & 15.34 & 1.67 & 17.09 \\
    \bottomrule
  \end{tabular}
  \label{table:full-core-power}
\end{table}

\FloatBarrier

\section{3-D Models \& Results}

%\subsection{3-D Model Setup}

%\subsection{3-D Numerical Results}

\section{Summary}

The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
diffusion method in cases where the neutron diffusion method provides