### 2024-09-11 verification.tex

\input{moltres}
\input{benchmark}

\section{MSRE Zero-Power Pump Experiments}

\section{Spalart-Allmaras Turbulence Model Verification}

\input{turbulence}