### 2024-10-06 verification.tex

Section \ref{sec:msr-vv} identifies salt flow-induced \gls{DNP} drift and the strong coupling
between neutronics and thermal-hydraulics as two critical multiphysics interactions that must be
accurately modeled in \gls{MSR} simulation software. This chapter describes Moltres and
subsequently covers several \gls{VV} studies for verifying and validating specific modeling
capabilities in Moltres which relate to the aforementioned multiphysics interactions.

Section \ref{sec:moltres-description} details the neutronics, thermal-hydraulics, and multiphysics
coupling schemes in Moltres for \gls{MSR} multiphysics simulations. Following the model
descriptions, the section also summarizes and assesses previous \gls{MSR} analyses
performed with Moltres. Section \ref{sec:cnrs-benchmark} presents verification results from Moltres
for the CNRS benchmark study. Section \ref{sec:msre-pump} presents key findings from a
collaborative \gls{VV} study based on the \gls{MSRE} pump transient tests conducted in the 1960s.
Section \ref{sec:turbulence} describes the Spalart-Allmaras turbulence model implementation in
Moltres and presents the model \gls{VV} results against common wall-bounded turbulent flow case
studies from literature.

\input{moltres}
\input{benchmark}

\section{MSRE Zero-Power Pump Experiments} \label{sec:msre-pump}

\input{msre-pump}

\section{Moltres Spalart-Allmaras Turbulence Model Verification} \label{sec:turbulence}

\input{turbulence}

\section{Summary}