### 2024-11-07 transient.tex
Chapter 6 intro

Chapter \ref{chap:msre} verified the hybrid $S_N$-diffusion method implemented in Moltres through
1-D, 2-D, and 3-D
$k$-eigenvalue simulations of the \gls{MSRE} against the OpenMC Monte Carlo neutron transport code
for molten salt reactor control rod modeling. The 3-D simulations validated the 3-D \gls{MSRE}
model in this work against \gls{MSRE} experimental data and the \gls[MSRE} numerical benchmark
study \cite{fratoni_molten_2020} in the context of the initial criticality and rod worth
measurement experiments. The 3-D simulations also showed that the hybrid method runs with
reasonable computational costs compared to the standard neutron diffusion method and exhibited
efficient strong scaling across compute nodes.

This chapter builds on the foundations of Chapter \ref{chap:msre} to verify and validate the hybrid
method for time-dependent 3-D simulations of reactivity-initiated transients. This chapter includes
two demonstrations using the hybrid method of the \gls{MSRE} rod drop and reactivity insertion
experiments. Beyond time-dependent modeling, the simulations will also verify and validate
compatibility between the existing neutronics/thermal-hydraulics coupling capabilities in Moltres
and the newly-implemented hybrid method.

Section \ref{sec:rod-drop}

\section{Rod Drop Experiment} \label{sec:rod-drop}

\subsection{Model Setup}

\subsection{Rod Drop Numerical Results \& Discussion}

\section{Reactivity Insertion Experiment} \label{sec:reactivity-insertion}

\subsection{Model Setup}

\subsection{Reactivity Insertion Numerical Results \& Discussion}

\section{Summary} \label{sec:transient-summary}