### 2024-11-13 main.tex

%% Package and Class "uiucthesis2021" for use with LaTeX2e.
\documentclass[11pt]{uiucthesis2021}
\doublespacing

\usepackage[acronym,toc]{glossaries}
\makeglossaries
\include{acros}

\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage[title,titletoc]{appendix}

\usepackage{placeins}
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{array}
\usepackage{microtype} % improves typography for PDF

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hhline}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\allowdisplaybreaks
\usepackage{color}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{xfrac}
\usepackage{bm}
\usepackage{stmaryrd}

\usepackage{threeparttable, tablefootnote}

\usepackage{environ}
\makeatletter

\usepackage{tabularx}
\usepackage{float}
\usepackage{enumitem}
\setlist[itemize]{noitemsep}
\usepackage{diagbox}
\usepackage{courier}
\usepackage{pdflscape}

\usepackage{cleveref}
\usepackage{datatool}
% \usepackage[numbers]{natbib}
\usepackage{notoccite}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[bibstyle=ieee,backend=bibtex8,sorting=none,eprint=false,isbn=false,doi=false]{biblatex}
\usepackage{textalpha}
\addbibresource{./bibliography.bib}
\AtEveryBibitem{%
  \clearfield{note}%
}

\usepackage{tikz}
\usepackage{tkz-euclide}
\usetikzlibrary{positioning, arrows, decorations, shapes}
\usetikzlibrary{shapes.geometric,arrows}
\definecolor{illiniblue}{HTML}{B1C6E2}
\definecolor{illiniorange}{HTML}{f8c2a2}
\definecolor{green}{HTML}{c2e2b1}
\tikzstyle{process} = [rectangle, rounded corners, thick, minimum width=7cm, minimum height=1cm, text centered, draw=black, fill=illiniblue, text width=18em]
\tikzstyle{object} = [ellipse, minimum width=2cm, thick, minimum height=2.2cm, text centered, draw=black, fill=illiniorange, text width=12em]
\tikzstyle{decision} = [diamond, thick, aspect=2, minimum width=2cm, minimum height=2cm, text centered, draw=black, fill=green, text width=6em]
\tikzstyle{arrow} = [thick,->,>=stealth]

\usepackage[ruled,linesnumbered,lined]{algorithm2e}
\SetArgSty{textrm}

%\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
\title{Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling}
\author{Sun Myung Park}
\department{Nuclear, Plasma \& Radiological Engineering}
\degreeyear{2024}
\committee{Associate Professor Kathryn D. Huff, Chair \& Co-Director of Research \\
           Assistant Professor Madicken Munk, Co-Director of Research, Oregon State University \\
           Professor Rizwan Uddin \\
           Professor Tomasz Kozlowski \\
           Professor Paul Fischer}

\newcounter{counterforappendices}

\begin{document}
\maketitle

\frontmatter
%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}

\input{abstract}

\end{abstract}
\addcontentsline{toc}{chapter}{Abstract}

%\begin{dedication}
%
%\end{dedication}

\begin{acknowledgments}

%\input{acks}

\end{acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

{
  \hypersetup{linkcolor=black}
  \tableofcontents
}

%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
\printglossary[title=List of Abbreviations,type=\acronymtype,nonumberlist,
nogroupskip=true]
%% Create a List of Symbols. The left column
%% is 0.7 inch wide and centered

\pagebreak
\mainmatter
\glsresetall

\chapter{Introduction}
\label{chap:intro}
\input{introduction}
\glsresetall

\chapter{Molten Salt Reactor Modeling and Simulation}
\label{chap:lit}
\input{lit}
\glsresetall

\chapter{Verification of Multiphysics Features in Moltres}
\label{chap:verification}
\input{verification}
\glsresetall

\chapter{Hybrid $S_N$-Diffusion Method}
\label{chap:hybrid}
\input{hybrid}
\glsresetall

\chapter{Demonstration of the Hybrid Method with the MSRE}
\label{chap:msre}
\input{msre}
\glsresetall

\chapter{Demonstration of Time-Dependent Reactivity-Initiated Transients}
\label{chap:transient}
\input{transient}
\glsresetall

\chapter{Conclusion}
\label{chap:conclusion}
\input{conclusion}
\glsresetall

\backmatter

{
\setstretch{1}
% \bibliographystyle{IEEEtran}
\printbibliography[heading=bibintoc,title={References}]
}

\clearpage
\setcounter{counterforappendices}{\value{page}}
\mainmatter
\setcounter{page}{\value{counterforappendices}}
\appendix
\input{appendix}

\end{document}
\endinput
%%