### 2024-12-04 presentation/msre-transients.tex

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Description of MSRE Pump Experiments}
  \begin{itemize}
    \item Pump start-up and coast-down transient flow-rate tests on the MSRE under zero-power
  conditions
    \item Starting from zero-power critical states with/without forced flow, the fuel salt pump
  was coasted down/started up
    \item Reactor was kept critical by a flux-servo controller moving the rod in response to flux
      changes
  \end{itemize}
  \begin{columns}
    \column[t]{5.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=.8\columnwidth]{images/msre-startup}
      \caption{MSRE start-up pump speed and flow rate experimental data
      \cite{prince_zero-power_1968}.}
    \end{figure}
    \column[t]{5.5cm}
    \begin{figure}
      \centering
      \includegraphics[width=.95\columnwidth]{images/msre-coastdown}
      \caption{MSRE coast-down pump speed and flow rate experimental data
      \cite{prince_zero-power_1968}.}
    \end{figure}
  \end{columns}
\end{frame}

%\begin{frame}
%  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
%  \begin{block}{\textbf{MSR Phenomena Involved}}
%    \begin{itemize}
%      \item DNP drift under time-varying flow
%      \item Loss of delayed neutrons due to out-of-core DNP decay
%    \end{itemize}
%  \end{block}
%  \begin{block}{\textbf{Aim of the Study}}
%    Reproduce the reactivity curve of the control rod response with a 2-D axisymmetric model of the
%    MSRE in Moltres.
%  \end{block}
%  \begin{block}{\textbf{Study Objectives}}
%    \begin{itemize}
%      \item Develop a V\&V study based on the MSRE pump start-up \& coast-down
%        transients that is easily reproducible
%      \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
%        (formerly at Oregon State University)
%      \item Validate Moltres against MSRE experimental data for modeling \gls{DNP} drift
%    \end{itemize}
%  \end{block}
%\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \begin{columns}
    \column{6.5cm}
    \begin{block}{\textbf{MSR Phenomena Involved}}
      \begin{itemize}
        \item DNP drift under time-varying flow
        \item Loss of delayed neutrons due to out-of-core DNP decay
      \end{itemize}
    \end{block}
    \begin{block}{\textbf{Aim of the Study}}
      \begin{itemize}
         \item Develop a V\&V study based on the MSRE pump start-up \& coast-down
           transients that is easily reproducible
         \item Verify Moltres against QuasiMolto in collaboration with Aaron Reynolds
           (formerly at Oregon State University)
         \item Validate Moltres against MSRE experimental data for modeling \gls{DNP} drift
       \end{itemize}
    \end{block}
    \begin{block}{\textbf{Modeling Approach}}
      \begin{itemize}
        \item Two neutron energy groups and six DNP groups
        \item Group constants generated using a representative unit cell model in SCALE
        \item Ignored upper and lower plena in the geometry
        \item Looped salt flow with a salt circulation time of 25.2 s
        \item $k$-eigenvalue calculation at every time-step in place of reactivity control
        \item Timestep size of 0.1 s
      \end{itemize}
    \end{block}
    \column{4.5cm}
    \begin{figure}[htb]
      \centering
      \includegraphics[width=0.8\columnwidth]{images/msre-2d}
      \caption{2-D axisymmetric model of the \gls{MSRE}.The gray and white regions represent fuel salt
      and graphite, respectively. Fuel salt channels are 1.25-cm wide and spaced at 5-cm intervals.}
      \label{fig:pump-geom}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Static \& Steady-Flow Conditions}
  \begin{table}[htb]
    \centering
    \caption{Multiplication factors $k_\text{eff}$ under static and steady salt flow conditions, and
    reactivity changes $\Delta\rho$ due to \gls{DNP} flow from the Moltres and QuasiMolto \gls{MSRE}
    models.}
    \begin{tabular}{l S S S}
      \toprule
      \multirow{2}{*}{Code} & \multicolumn{2}{c}{$k_\text{eff}$} & {$\Delta\rho$ due to} \\
                            & {Static} & {Steady flow} & {\gls{DNP} flow} \\
                            \cmidrule(r){1-1} \cmidrule(rl){2-3} \cmidrule(l){4-4}
      Moltres & 1.05625 & 1.05311 & -282.4 \\
      QuasiMolto & 1.05603 & 1.05290 & -281.7 \\
      \bottomrule
    \end{tabular}
    \label{table:msre-pump-keff}
  \end{table}

  \begin{itemize}
    \item Excellent agreement between Moltres and QuasiMolto under static and steady-flow conditions prior
  to the pump transients.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Neutron Flux Distributions Under Static Conditions}
  \begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{centerline_flux}
      \caption{Normalized centerline neutron fluxes.}
      \label{fig:centerline-flux-dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{midplane_flux}
      \caption{Normalized midplane neutron fluxes.}
      \label{fig:midplane-flux-dist}
    \end{subfigure}
    \caption{Neutron flux distributions comparing QuasiMolto and Moltres
    \gls{MSRE} models under static conditions. Maximum relative difference $\approx 0.8 \%$.}
    \label{fig:centerline-flux}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{DNP Distributions Under Static Conditions}
  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{centerline_pre}
      \caption{Normalized centerline \gls{DNP} distributions.}
      \label{fig:centerline-pre-dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\columnwidth}
      \centering
      \includegraphics[width=\columnwidth]{midplane_pre}
      \caption{Normalized midplane \gls{DNP} distributions.}
      \label{fig:midplane-pre-dist}
    \end{subfigure}
    \caption{\gls{DNP} distribution from Moltres and ratios comparing QuasiMolto and
    Moltres models under static \gls{MSRE} conditions. Maximum relative difference $\approx 0.1 \%$.}
    \label{fig:centerline-pre}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Start-Up Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    start-up transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Observations}
  \begin{itemize}
    \item Steep initial reactivity loss due to DNP outflow
    \item Reactivity loss plateaus temporarily before dropping due to DNP reentry
    \item Circulation of the excess DNP cause reactivity oscillations
    \item Oscillations eventually dissipate due to decay of excess DNP
    \item Excellent agreement between Moltres and QuasiMolto
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Start-Up Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{start-up-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    start-up transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Discrepancies against MSRE Data}
  \begin{itemize}
    \item MSRE reported a smaller final reactivity loss
    \item MSRE did not report a reactivity plateau (due to DNP present in the upper and lower plena)
    \item MSRE did not log significant reactivity oscillations after the first peak
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Pump Coast-Down Transient Results}
  \begin{columns}
  \column{5.5cm}
  \begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{coast-down-v2-reactivity}
    \caption{Reactivity loss (relative to static no-flow conditions) during the \gls{MSRE} pump
    coast-down transient.}
    \label{fig:start-up-reactivity}
  \end{figure}
  \column{5.5cm}
  \textbf{Observations}
  \begin{itemize}
    \item Reactivity loss remains briefly stable at the start due to flow inertia
    \item Monotonic decline in reactivity loss over time
  \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{V\&V Study 2: MSRE Pump Start-up \& Coast-Down Transients}
  \textbf{Study Outcome}
  \begin{itemize}
    \item Moltres and QuasiMolto are highly consistent with each other in their looped DNP flow
      modeling capabilities
    \item Moltres and QuasiMolto reproduced key reactivity behavior in response to varying salt
      flow velocities
    \item Some discrepancies observed between the numerical and experimental results due to the
      omission of upper and lower plena
    \item This study can be used for verifying advection capabilities in other MSR simulation tools
    \item The distinct oscillations serve as valuable reference points to facilitate code-to-code
      comparisons
  \end{itemize}
\end{frame}