### 2024-12-04 presentation/objectives.tex

\begin{frame}
  \frametitle{Objectives}
  \begin{block}{\textbf{Overarching Goal of this Work}}
    \textbf{Improve Moltres as a MSR simulation tool by enabling a wider range of MSR simulation
    scenarios.}
  \end{block}
  \begin{block}{\textbf{Objectives}}
    \begin{enumerate}
      \item \textbf{Improve Multiphysics Capabilities in Moltres}
      \begin{itemize}
        \item \textbf{Verify and Validate Existing Multiphysics Coupling Capabilities in Moltres}
        \item \textbf{Implement a RANS-based Turbulence Model in Moltres}
      \end{itemize}
      \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
    \end{enumerate}
  \end{block}
\end{frame}