### 2024-12-09 presentation/objectives.tex

\begin{frame}
  \frametitle{Objectives}
  \begin{block}{\textbf{Overarching Goal of this Work}}
    Improve Moltres as a MSR simulation tool by enabling a wider range of MSR simulation
    scenarios.
  \end{block}
  \begin{block}{\textbf{Objectives}}
    \begin{enumerate}
      \item \textbf{Improve Multiphysics Capabilities in Moltres}
      \begin{itemize}
        \item Verify and validate existing multiphysics coupling capabilities in Moltres
        \item Implement a RANS-based turbulence model in Moltres
      \end{itemize}
      \item \textbf{Develop a Hybrid Neutronics Method for Control Rod Modeling in Moltres}
      \begin{itemize}
        \item Develop and implement the hybrid $S_N$-diffusion method in Moltres
        \item Verify the hybrid method against high-fidelity neutronics codes and MSRE
          experimental data
        \item Demonstrate a time-dependent reactivity-initiated simulation using the hybrid
          method
      \end{itemize}
    \end{enumerate}
  \end{block}
\end{frame}