### 2024-12-09 presentation/turb-motivation.tex

\begin{frame}
  \frametitle{Motivation for Turbulence Modeling Implementation}
  \textbf{Turbulent Flows in MSRs}
  \begin{itemize}
    \item The Reynolds number of salt flow in MSRs can go up to $10^6$ in the MSFR core
      \cite{fiorina_modelling_2014}.
    \item Turbulent effects expected in MSRs
    \begin{itemize}
      \item Turbulent diffusion of heat and \gls{DNP}
      \item Flow separation and recirculation zones due to abrupt geometry changes
    \end{itemize}
    \item Temperature hotspots may induce thermal stress on structural
      components and temperature-induced reactivity effects.
  \end{itemize}
  \pause
  \begin{block}{\textbf{Area of Improvement for Moltres for MSR Modeling}}
    Moltres does not currently have turbulence modeling capability for simulating
    turbulent flows in MSRs.
  \end{block}
\end{frame}